\newpage

\section{Research Summary}

My early work from 1989 through around 2006 focused primarily on parallel and distributed computing systems, including work on experimental language designs and compilers for parallel systems based on dataflow architecture. I conducted this work while a graduate student at the Illinois Institute of Technology and in graduate and postgraduate work at Argonne National Laboratory.
\vspace{5pt}

Most of these works are focused on messaging middleware to support cluster computing, grid computing, and scalable I/O. In all cases, my research was accompanied by working software to support the research being described in my research papers. I often found myself as one of the people who not only could implement the systems but who understood software engineering methods well enough to produce designs/implementations of high quality. This has resulted in a recent shift to focus more energy on software engineering, especially to address software created in research domains (also known as research software).
\vspace{5pt}

While high-performance computing still plays a role in what I do, since becoming tenured and reaching full-professor statues, I have been increasingly drawn to my interdisciplinary interests. I have many collaborations within and outside of the computing discipline. Being trained in the liberal arts and sciences, I have interests in many subjects, and collaborations in these disciplines have resulted in many publications, some of which might initially appear strange but are part of a coherent narrative and worldview. At the time of writing, I have collaborated on works in Anthropology, Biology, English, History, and Engineering domains. One of these collaborations (with English) resulted in significant scholarly/research output in digital humanities, which resulted further in the creation of a digital humanities program at my university (with me as co-founder and the inaugural co-director).
\vspace{5pt}

The rest is organized as a brief summary of each research area and selected publications. The most recent (and funded) works appear first.


\input{rd-cam2-se}
\input{rd-fossil-record}
\input{rd-se4science}
\input{rd-hoc}
\input{rd-filesystems}
\input{rd-digital-humanities}
\input{rd-message-passing}
\input{rd-citep}



